---
title: "Assignment 3"
author: "Marco Palombo"
date: "4/19/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
```

```{r}
# import 
data <-  (data <- read_csv(here("HW3_data.csv"))) %>% 
 clean_names() %>% 
  mutate(q_agg_kwh = q_high_kwh + q_low_kwh)
  
# given
scc_co2_cost <- 51 #$/metric ton
el_price_cents <- 10 #cents/kWh

# conversion factors
tonne_to_lbs = 2204.62 # pounds/metric ton
```

```{r}
# linear fits to the high and low income data
high_fit <- lm(q_high_kwh ~ price_cents, data = data)
low_fit <- lm(q_low_kwh ~ price_cents, data = data)

# now create functions based on parameters from these fits
high_func_q <- function(price){  
  quantity = high_fit$coefficients[2]*price + high_fit$coefficients[1]
}

low_func_q <- function(price){  
  quantity = low_fit$coefficients[2]*price + low_fit$coefficients[1]
}

# inverse functions, use same constants to ensure inverse
high_func_p <- function(quantity){  
  price = (quantity - high_fit$coefficients[1])/high_fit$coefficients[2]
}

low_func_p <- function(quantity){  
  price = (quantity - low_fit$coefficients[1])/low_fit$coefficients[2]
}

```

1. One kWh of electricity emits 0.85 pounds of CO2. Assuming that the interim SCC correctly
reflects the total social cost of one metric ton of CO2, what is the marginal externality cost
per kwH of electricity?

```{r}
# part one calculation
mec_el <- 0.85/tonne_to_lbs*scc_co2_cost
mec_el_cents <- round(mec_el*100,2)

```

Use dimensional analysis for solution:
$\mathrm{\frac{0.85~lbs}{kWh}*\frac{tonne~CO_2}{2204.62~lbs~CO_2}*\frac{\$51}{1~tonne~CO_2}*\frac{100~cents}{1~\$}}$ = `r round(mec_el_cents,2)`$\mathrm{\frac{cents}{kWh}}$

2. What is the aggregate monthly demand curve for electricity? What is the supply curve for
electricity? What is the “benefit” to consumers under the status quo? What is the “benefit”
to producers under the status quo? What is the environmental cost under the status quo?

```{r}
#linear fit to the aggregate vector
agg_fit <- lm(q_agg_kwh ~ price_cents, data = data)

# aggregate function price as input simply sum of coefficients 
agg_func_q <- function(price){
  quantity = agg_fit$coefficients[2]*price + agg_fit$coefficients[1]
  }

agg_func_p <- function(quantity){  
  price = (quantity - agg_fit$coefficients[1])/agg_fit$coefficients[2]
}
```

```{r}
#plug in 10 cent equilibrium price to determine quantity supplied at this price
eq_quantity <- round(agg_func(el_price_cents),0)

# use this point and given zero intercept to determine slope first having quantity be function of cost
supply_slope <- eq_quantity/el_price_cents
# now create a function of quantity as function of price

supply_func <- function(price) {
  quantity = supply_slope*price
}

supply_func_inv <- function(quantity){
  price = quantity/supply_slope
}
```



```{r}
#calculate excesses
cons_exc <- integrate(agg_func_inv, 0, eq_quantity)$value  - integrate(supply_func_inv, 0, eq_quantity)$value

prod_exc <- integrate(supply_func_inv, 0, eq_quantity)$value

env_cost <- eq_quantity*mec_el_cents

```
Aggregate monthly demand curve is horizontal sum of high and low income monthly demand curves. Start by creating linear fit of quantity demanded as function of price for both income groups. Given these are in the same format coefficients can be directly summed.

High income monthly demand function:\
high_income_demand(price) = `r high_fit$coefficients[2]` $\mathrm{\frac{kWh}{\$}}$ * price(cents) + `r high_fit$coefficients[1]` kWh

Low income monthly demand function:\
low_income_demand(price) = `r low_fit$coefficients[2]`$\mathrm{\frac{kWh}{\$}}$ * price(cents) + `r low_fit$coefficients[1]` kWh

Aggregate monthly demand function:\
aggregate_demand(price)= `r high_fit$coefficients[2]+low_fit$coefficients[2]`$\mathrm{\frac{kWh}{cent}}$ * price(cents) + `r high_fit$coefficients[1]+low_fit$coefficients[1]` kWh

To get our supply curve we can use our current price and the given fact of a zero intercept, along with assumption of linearity, to generate a line. Plug in given $0.10 equilibrium price into aggregate demand curve to determine quantity supplied at this price. Next generate equation for a line that intersects this point and the origin. In this case I used cost as the independent variable, if I was to use quantity instead the linear factor (slope) would simply inverted. 

Supply(price) = `r supply_slope` $\mathrm{{kwh}{cent}}$ *cost

Next I will assume the term "benefit" is excess as defined earlier in this course.

Consumer "benefit"(excess):
Area under aggregate demand curve between 0 and `r eq_quantity` kWh minus area under supply curve from zero to `r eq_quantity`.
`r cons_exc` cents

Supplier "benefit"(excess):
Area under supply curve from zero to `r eq_quantity`.
`r prod_exc` cents

Environmental cost:
Area under MEC curve between 0 and `r eq_quantity`. Given MEC curve is a constant this will simply be the product of MEC value and `r eq_quantity`.
`r env_cost` cents

3.  How is the current consumer benefit divided between “high” and “low” income consumers?

Consumer benefit is equal to the sum of the consumer benefit of high income consumers and the consumer benefit of low income consumers. In order to find each of these the first step is to find the intercept with the supply curve for both high and low income consumers. 

```{r}
# find intercept in quantity space
low_int_quantity <- round(uniroot(function(quantity) supply_func_inv(quantity)-low_func_inv(quantity), c(0,400000), tol = .01)$root,0)

high_int_quantity <- round(uniroot(function(quantity) supply_func_inv(quantity)-high_func_inv(quantity), c(0,600000), tol = .01)$root,0)

high_exc <- integrate(high_func_inv, 0, high_int_quantity)$value  - integrate(supply_func_inv, 0, high_int_quantity)$value

low_exc <- integrate(low_func_inv, 0, low_int_quantity)$value  - integrate(supply_func_inv, 0, low_int_quantity)$value

test = cons_exc-high_exc-low_exc
```

```{r}
#sanity check
data_check <- data %>% 
  mutate(high_est = high_func(price_cents),
         low_est = low_func(price_cents),
         cost_est_high = high_func_inv(q_high_kwh),
         cost_est_low = low_func_inv(q_low_kwh),
         agg_est = agg_func(price_cents),
         sum = high_est + low_est,
         supply = supply_func(price_cents))

ggplot(data = data_check)+
  geom_line(aes(x = high_est, y = price_cents), color = "black")+
  geom_line(aes(x = q_high_kwh, y = price_cents), color = "black")+
  geom_line(aes(x = low_est, y = price_cents), color = "blue")+
  geom_line(aes(x = q_low_kwh, y = price_cents), color = "red")+
  geom_line(aes(x = supply, y = price_cents), color = "blue")+
  geom_line(aes(x = agg_est, y = price_cents), color = "cyan")+
  geom_hline(yintercept = mec_el_cents, color = "orange")+
  geom_point(aes(x = high_int_quantity, y = high_func_inv(high_int_quantity)))+
  geom_point(aes(x = low_int_quantity, y = low_func_inv(low_int_quantity)))

```
Consumer "benefit"(excess):
Area under aggregate demand curve between 0 and `r eq_quantity` kWh minus area under supply curve from zero to `r eq_quantity`.
`r cons_exc` cents





