---
title: "Assignment 3"
author: "Marco Palombo"
date: "4/19/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(scales)
library(equatiomatic)
```
## Overview

```{r}
# Read in the main data set
carbon <- read_csv(here("HW3_data.csv")) %>% 
  clean_names() %>% 
  select(-x1)
```

### 1

One kWh of electricity emits 0.85 pounds of CO2. Assuming that the interim SCC correctly reflects the total social cost of one metric ton of CO2, what is the marginal externality cost per kwH of electricity?

Assuming that one kWh of electricity emits 0.85 pounds of CO2 and that the interim SCC of $51/ton is accurate, this section calculates the MEC per kWh.

```{r}
# Create a longer dataset where income is a variable, not a column 
carbon_long <- carbon %>% 
  pivot_longer(cols = c(q_low_kwh, q_high_kwh),
               names_to = 'income_level',
               values_to = 'kwh') %>% 
  mutate(income_level = case_when(income_level == 'q_low_kwh' ~ 'low',
                   income_level == 'q_high_kwh' ~ 'high'))

# solve for MEC
price_ton <- 51
#2205 lbs in a metric ton. Divide by 0.85 to get kwh produced per ton
kwh_ton <- 2204.6/0.85
# Multiply by Biden price to get MEC/kWh (Basically, $51 is the price for 2594 kwh)
interim_price_kwh <- price_ton/kwh_ton
```

Marginal externality cost per kWh of electricity is $`r round(interim_price_kwh, 4)`.

### 2
What is the aggregate monthly demand curve for electricity? What is the supply curve for  electricity? What is the “benefit” to consumers under the status quo? What is the “benefit” to producers under the status quo? What is the environmental cost under the status quo?

Low income demand curve:
```{r}
# linear fit low income data
demand_low <- lm(price_cents ~ kwh, income_level=='low', 
                 data = carbon_long) 
extract_eq(model = demand_low, use_coefs = TRUE, coef_digits = 5)
```

High income demand curve:
```{r}
# linear fit to high income data
demand_high <- lm(price_cents ~ kwh, income_level=='high',
                  data = carbon_long) 
extract_eq(model = demand_high, use_coefs = TRUE, coef_digits = 5)

# Generalized demand model from sample code 
demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
# Generalized aggregate demand from code sample
demand_agg <- function(p){
  q <- demand(p, demand_low) + demand(p, demand_high)
  return(q)
}

```

Perform horizontal sum of low and high income demand curves to get aggregate demand curve. This does not follow a perfect line due to the kink at the top. 

```{r}
# Make a vector and extract the lm 
price = seq(0, 30, length.out = 100)
Qagg <- map(price, demand_agg) %>% 
  unlist()
agg_df<- tibble(Qagg = Qagg, price = price)
```

To calculate the supply curve plug in given 10 cent equilibrium price and use the resulting quantity combined with given zero intercept to create function for line intersecting these two points.

```{r}
kwh_agg <- demand_agg(10)
#536,719.5 kWh consumed @ $0.10
supply_slope <- 10/kwh_agg
supply <- function(p){
  q <- p/supply_slope
  return(q)
}
```

P = 0.0000186Q

#### Market for Electricity
```{r}
ggplot(agg_df, aes(Qagg, price)) +
  geom_line(color = "red") +
  annotate("text", x = 400000, y = 17, 
           label = "Aggregate Demand", angle = -30) +
  geom_abline(color = "blue",
              intercept = demand_high$coefficients[1],
              slope= demand_high$coefficients[2]) +
  annotate("text", x = 300000, y = 15, 
           label = "High-Income Demand", angle = -38) +
  geom_abline(color = "green",
              intercept = demand_low$coefficients[1],
              slope= demand_low$coefficients[2]) +
  annotate("text", x = 150000, y = 10, 
           label = "Low-Income Demand", angle = -55) +
  geom_abline(color = "orange",
              intercept = 0,
              slope= supply_slope) +
  annotate("text", x = 300000, y = 7, 
           label = "Supply", angle = 17) +
  labs(x = "Electricity Consumed (kWh)",
       y = "Price (cents/kWh)")
  
```


Now that we have curves we can calculate excesses, which is caled benefit in this excercise for reasons which are not clear to me. 

```{r}
#consumers
cs_baseline <- 0.5*kwh_agg*(0.305 - 0.10)
#producers
ps_baseline <- 0.5*kwh_agg*0.10
# env cost
env_cost_baseline <- interim_price_kwh*kwh_agg
# $10,553.75
```

- Consumer Benefit: Area under aggregate demand curve and above equilibrium price between zero and quantity demanded at equilibrium price
$`r round(cs_baseline, 0)`
- Producer Benefit: Area under equilibrium price and above supply curve between zero and quantity demanded at equilibrium price
$`r round(ps_baseline, 0)`
- Environmental Cost: MEC multiplied by quantity demandedat equilibrium price
$`r round(env_cost_baseline, 0)`


### 3
 
How is the current consumer benefit divided between “high” and “low” income consumers?
 
Use the current price to calculate the amount each group will consume at this level. Then repeat consumer benefit calculation for both groups considering their respective purchased quantities and demand curves.

```{r}
# low income
cs_low <- 0.5*(.2337 - .10)*demand(10, demand_low)
# high income
cs_high <- 0.5*(.3161-.10)*demand(10, demand_high)
```

- Low-income CS: 
$`r round(cs_low, 0)`
- High-income CS: 
$`r round(cs_high, 0)`

### 4
Derive the optimal electricity tax (in cents per kWh) using the interim SCC. Noting that recent
research has shown the poor face a disproportionate share of the impacts from climate change,
assume that the climate externality is borne entirely by the “low” income group. What would
be the effects of this tax on:
(a) The amount of electricity produced and consumed
(b) The price of electricity
(c) Overall welfare of “high” income consumers
(d) Overall welfare of “low” income consumers

The "derivation" of optimal tax is simple. It will be equivalent to MEC calculated above.

Pretty impressive derivation huh?

It appears we are being asked to assume the entire environmental impact will be subtracvted from low income welfare, though I could be wrong, so that is what will be done.

```{r}
#define tax
tax_cents <- 1.97
# rewrite function including tax
demand_tax <- function(p, model){
  q <- (p - (model$coefficients[[1]]-1.97))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
} 
demand_agg_tax <- function(p){
  q <- demand_tax(p, demand_low) + demand_tax(p, demand_high)
  return(q)
}

# new demand curve intersects supply curve at 9.32 cents this will be out new price
p_new <- 9.32
# new quantity
q_new <- demand_agg_tax(p_new)
```

plug in the new price and quantity to get updated values for excesses.

```{r}
#price and demand changes
demand_change <- kwh_agg - q_new
price_change <- 10 - p_new

# new demand high income consumers
demand_high_tax <- demand_tax(p_new, demand_high)
#new consumer surplus
cs_high_tax <- (0.5 * (demand_high$coefficients[1]-tax_cents-p_new) * demand_high_tax)/100
#high income consumers    
welfare_high_tax <- cs_high_tax - (demand_high_tax * 0.0197)
#changes
welfare_high_change <- cs_high - welfare_high_tax

#assume that environmental damage affects low income consumers in no tax scenario
welfare_low_baseline <- cs_low - env_cost_baseline
# new low income demand
demand_low_tax <- demand_tax(p_new, demand_low)
#environmetal cost at new demand levels
env_cost_tax <- interim_price_kwh * q_new
# suprlus
cs_low_tax <- (0.5 * (demand_low$coefficients[1]-tax_cents-p_new) * demand_low_tax)/100
# subtract taxes and environmental damage to get welfare
welfare_low_tax <- cs_low_tax - (demand_low_tax * 0.0197) - env_cost_tax
welfare_low_change <- welfare_low_baseline - welfare_low_tax

#producer surplus
ps_tax <- 0.5*q_new*0.0932
ps_change <- ps_baseline - ps_tax

#environmetal damages
damage_change <- env_cost_baseline - env_cost_tax

# tax revenue
tax_revenue <- interim_price_kwh*q_new

```


A: 
Electricity produced/consumed decreases by `r round(demand_change, 0)`kWh. New total consumption is `r round(q_new, 0)`

B: 
The price of electricity decreases by $`r round(price_change, 2)`.
New price is $`r p_new`/kWh. 

C: 
The welfare of high-income consumers decreases by $`r round(welfare_high_change, 0)`.

D: 
The welfare of low-income consumers decreases by $`r round(welfare_low_change, 0)`. 

Note: this is becasue we applied the negative environmental externality to the original case as well. Now they are experiencing a slightly lower environmental externality but are also now taxed.

E: 
Producer surplus decreases by $ `r round(ps_change, 0)`.

F: 
Environmental damage*decreases by $`r round(damage_change, 0)`.

G: 
Tax revenue is $`r round(tax_revenue, 0)`.
Note: this is equal to the total environmental cost.

### 5 

```{r}
# repeat calcs for each hypothetical SCC value
tax_75 <- 2.89
tax_100 <- 3.86
tax_125 <- 4.82
tax_150 <- 5.78

# Put into vector 
scc_vec <- c(tax_75, tax_100, tax_125, tax_150)

# pre-tax consumption 
low_share <- demand(10, demand_low)/kwh_agg
high_share <- demand(10, demand_high)/kwh_agg



```

The following are the high-income welfare, low-income welfare, and producer surplus at five SCC levels:

```{r}
# $51 
low_redist_51 <- low_share*tax_revenue
low_welfare_51 <- low_redist_51 + welfare_low_tax
#New welfare is equal to initial welfare --- (-$2422)
high_redist_51 <- high_share*tax_revenue
# Get $7612 back 
high_welfare_51 <- high_redist_51 + welfare_high_tax
```

```{r}
#$75
demand_scc <- function(p, model, scc){
  q <- (p - (model$coefficients[[1]] - scc_vec[scc]))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
} 
demand_agg_75 <- function(p){
  q <- demand_scc(p, demand_low, 1) + demand_scc(p, demand_high, 1)
  return(q)
}
p_75 <- uniroot(function(p)
  demand_agg_75(p) - supply(p),
  interval = c(0,20))
#Gives new P of 9.002
demand_75 <- demand_agg_75(p_75$root)
revenue_damage_75 <- demand_75*.0289
#Low Income @ $75
demand_low_75 <- demand_scc(p_75$root, demand_low, 1)
cs_low_75 <- (0.5*demand_low_75*(demand_low$coefficients[1]-tax_75-p_75$root))/100
# Welfare is CS - tax + money back 
welfare_low_75 <- cs_low_75 - (demand_low_75 * .0289) - revenue_damage_75 + low_share*revenue_damage_75
              
# High Income @$75
demand_high_75 <- demand_scc(p_75$root, demand_high, 1)
cs_high_75 <- (0.5*demand_high_75*(demand_high$coefficients[1]-tax_75-p_75$root))/100
welfare_high_75 <- cs_high_75 - (demand_high_75*0.0289) + high_share*revenue_damage_75
# Producer @ $75
ps_75 <- 0.5*demand_75*(p_75$root/100)
```

```{r}
#$100
demand_agg_100 <- function(p){
  q <- demand_scc(p, demand_low, 2) + demand_scc(p, demand_high, 2)
  return(q)
}
p_100 <- uniroot(function(p)
  demand_agg_100(p) - supply(p),
  interval = c(0,20))
#Gives new P of 8.667
demand_100 <- demand_agg_100(p_100$root)
revenue_damage_100 <- demand_100*.0386
#Low Income @ $100
demand_low_100 <- demand_scc(p_100$root, demand_low, 2)
cs_low_100 <- (0.5*demand_low_100*(demand_low$coefficients[1]-tax_100-p_100$root))/100
welfare_low_100 <- cs_low_100 - (demand_low_100*0.0386) - revenue_damage_100 + low_share*revenue_damage_100
              
# High Income @$100
demand_high_100 <- demand_scc(p_100$root, demand_high, 2)
cs_high_100 <- (0.5*demand_high_100*(demand_high$coefficients[1]-tax_100-p_100$root))/100
welfare_high_100 <- cs_high_100 - (demand_high_100*0.0386) + high_share*revenue_damage_100
# Producer @ $100
ps_100 <- 0.5*demand_100*(p_100$root/100)
```

```{r}
#$125
demand_agg_125 <- function(p){
  q <- demand_scc(p, demand_low, 3) + demand_scc(p, demand_high, 3)
  return(q)
}
p_125 <- uniroot(function(p)
  demand_agg_125(p) - supply(p),
  interval = c(0,20))
#Gives new P of 8.336
demand_125 <- demand_agg_125(p_125$root)
revenue_damage_125 <- demand_125*.0482
#Low Income @ $125
demand_low_125 <- demand_scc(p_125$root, demand_low, 3)
cs_low_125 <- (0.5*demand_low_125*(demand_low$coefficients[1]-tax_125-p_125$root))/100
welfare_low_125 <- cs_low_125 - (demand_low_125*0.0482) - revenue_damage_125 + low_share*revenue_damage_125
              
# High Income @$125
demand_high_125 <- demand_scc(p_125$root, demand_high, 3)
cs_high_125 <- (0.5*demand_high_125*(demand_high$coefficients[1]-tax_125-p_125$root))/100
welfare_high_125 <- cs_high_125 - (demand_high_125*0.0482) + high_share*revenue_damage_125
# Producer @ $125
ps_125 <- 0.5*demand_125*(p_125$root/100)
```

```{r}
#$150
demand_agg_150 <- function(p){
  q <- demand_scc(p, demand_low, 4) + demand_scc(p, demand_high, 4)
  return(q)
}
p_150 <- uniroot(function(p)
  demand_agg_150(p) - supply(p),
  interval = c(0,20))
# New P of $8.0045, but actual price paid is P + tax 
demand_150 <- demand_agg_150(p_150$root)
revenue_damage_150 <- demand_150*.0578
# Low income @$150
demand_low_150 <- demand_scc(p_150$root, demand_low, 4)
cs_low_150 <- (0.5*demand_low_150*(demand_low$coefficients[1]-tax_150-p_150$root))/100
welfare_low_150 <- cs_low_150 - (demand_low_150*0.0578) - revenue_damage_150 + low_share*revenue_damage_150
              
# High Income @$150
demand_high_150 <- demand_scc(p_150$root, demand_high, 4)
cs_high_150 <- (0.5*demand_high_150*(demand_high$coefficients[1]-tax_150-p_150$root))/100
welfare_high_150 <- cs_high_150 - (demand_high_150*0.0578) + high_share*revenue_damage_150
# Producer @ $100
ps_150 <- 0.5*demand_150*(p_150$root/100)
```

A: High-income welfare

- 51 USD/ton: \
$`r round(high_welfare_51, 0)`. \
- 75 USD/ton: \
$`r round(welfare_high_75, 0)`. \
- 100 USD/ton: \
$`r round(welfare_high_100, 0)`.\
- 125 USD/ton: \
$`r round(welfare_high_125, 0)`.\
- 150 USD/ton: \
$`r round(welfare_high_150, 0)`.\

High income welfare decreases with increasing SCC levels assuming taxes are deducted from consumer surplus and the redistribution is added based on their pre-tax share of electricity consumption (`r round(high_share, 2)*100`%).

B: Low-income welfare 

- SCC of 51 USD/ton: 
$`r round(low_welfare_51, 0)` \
- SCC of 75 USD/ton: \
$`r round(welfare_low_75, 0)` \
- SCC of 100 USD/ton: \
$`r round(welfare_low_100, 0)`\
- SCC of 125 USD/ton: \
$`r round(welfare_low_125, 0)`\
- SCC of 150 USD/ton: \
$`r round(welfare_low_150, 0)`\

Low income welfare decreases with increasing SCC levels assuming taxes are deducted from consumer surplus, the entire environmental cost is deducted, and the redistribution is added based on their pre-tax share of electricity consumption (`r round(low_share, 2)*100`%).

C: Producer Welfare

- SCC of 51 USD/ton: \
$`r round(ps_tax, 0)`\ 
- SCC of 75 USD/ton: \
$`r round(ps_75, 0)` \
- SCC of 100 USD/ton: \
$`r round(ps_100, 0)`\
- SCC of 125 USD/ton: \
$`r round(ps_125, 0)`\
- SCC of 150 USD/ton: \
$`r round(ps_150, 0)`\

Producer surplus decreases as SCC increases due to declining quantities consumed.


### 6
Suppose the “high” income group has access to expensive home solar generation. This lowers
the electricity demand curve for the “high” income group by half (vertically). Under this new
demand:
(a) What is total electricity consumption?
(b) What is the total environmental externality?
(c) What value of the electricity tax makes the total environmental damage the same as the
damage when solar panels are available to the high income group?
 

```{r}
# New high-income demand 
demand_high_solar <- function(p){
  q <- (p - (demand_high$coefficients[[1]]-1.97)/2)/demand_high$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
#Low income demand, again 
demand_low_solar <- function(p){
  q <- (p - (demand_low$coefficients[[1]]-1.97))/demand_low$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
demand_agg_solar <- function(p){
  q <- demand_high_solar(p) + demand_low_solar(p)
  return(q)
}
p_solar <- uniroot(function(p)
  demand_agg_solar(p) - supply(p),
        interval = c(0,20))
price_solar <- p_solar$root
# New price is 5.845
demand_solar <- demand_agg_solar(price_solar)
#313691 is new demand 
env_cost_solar <- demand_solar*0.0197
# Need SCC equation when it is equal to solar equation in terms of damage 
tax_final <- (.0197/(env_cost_solar/env_cost_tax))
```

A: Total electricity consumption with solar is `r round(demand_solar, 0)` kWh. 

B: Total environmental cost is $`r round(env_cost_solar, 0)`.

C: The tax to make the environmental damage associated with the 51 USD SCC and no solar equal to the environmental damage with solar would be $`r round(tax_final, 4)`.

Note: this is roughly double the tax initially set based on the MEC at an SCC of $51